### 2022-09-23 manuscript/Chapter3-manuscript.Rmd
update methods MS

---
title: "Hunting experience and prey variability jointly shape individual foraging specialisation in a predator-prey videogame"
output: 
  word_document:
    reference_docx: Chapter3-styles.docx
    fig_caption: yes
bibliography: references.bib
csl: e-letters.csl
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(data.table)
library(brms)

```

```{r}

path <- file.path(here::here(), "data")
dat <- fread(file.path(path, "FraserFrancoetalXXXX-data.csv"),
             select = c("match_encode_id", "game_mode",
                        "realm_id", "environment_id", "predator_avatar_id",
                        "predator_id", "pred_game_duration",
                        "total_xp_killer", "cumul_xp_killer", "hunting_success",
                        "pred_speed", "prey_avg_speed"))

dat <- unique(dat)
dat[, c(1:6) := lapply(.SD, function (x) as.factor(x)), .SDcols = c(1:6)]

```

\newpage
# ABSTRACT {.unnumbered}

Keywords: foraging behavior, individual specialization, experience, learning, prey predictability, online videogames

\newpage
# INTRODUCTION {.unnumbered}

Individual variation in predator foraging behavior is increasingly recognized as a major driver of trophic interactions and community dynamics [@Griffen.etal2012;@Michalko.Pekar2016;@Moran.etal2017;@Michalko.etal2021]. Indeed, predator populations are often composed of assemblages of individuals specializing in different foraging strategies or resources irrespective of sexual, morphological, or age-related differences [@Estes.etal2003;@Tinker.etal2008;@kernaleguenDeterminantsIndividualForaging2015;@Phillips.etal2017]. A growing body of evidence suggests that ecological interactions, such as predator-prey interactions, can be major drivers of such individual foraging specialization [@Araujo.etal2011;@toscanoPersonalityForagingBehavior2016]. When they hunt, predators often use techniques that are fine-tuned to the type of prey that they encounter [@Davoren.etal2003;@Estes.etal2003;@Woo.etal2008;@Courbin.etal2018], and their capacity to use them effectively is contingent on periods of extensive practice (i.e. expertise). While the development of hunting expertise may be essential to maintain foraging success, we have have limited evidence for its role in predator foraging specialization, and the ecological/fitness consequences of such among individual behavioral differences for predator-prey interactions [@Dukas2019].

The integration of individual behavioral variation in the study of predator-prey interactions has gained traction in recent years, with empirical studies revealing important consequences for habitat use, functional responses, prey choice, and foraging rate [@Kobler.etal2009;@toscanoTraitmediatedFunctionalResponses2014;@Patrick.Weimerskirch2014;@Matsumura.Miyatake2022a]. However, an important and recurring challenge impeding research on predator-prey behavioral interactions, at the individual level, is the need of collecting data simultaneously on both the predator and prey. We have recently demonstrated with behavioral data from an online predator-prey videogame that virtual systems can help in overcoming this challenge, and uncover subtle details on the mechanisms shaping these interactions [@FraserFranco.etal2022]. For instance, we tested the classical locomotor crossover hypothesis championed by @Huey.Pianka1981, and found support of its predictions only for active hunters, which concords with a recent experiment involving assassin bugs [@Matsumura.Miyatake2022a]. Thus, virtual systems may be useful in identifying general ecological patterns, which may in turn help in advancing the current gaps in predator-prey research. Other examples of virtual ecological studies show that predation regimes can drive individual variation in risk perception [@Beauchamp2020], that familiarity between prey has a positive indirect effect on survival [@Cere.etal2021], and that prey face contrasting natural and social selection regimes (Santostefano et al. in prep). Because virtual systems generate large volumes of data on prey and predators simultaneously throughout their lifetime in the game, they offer the opportunity to tackle fundamental questions about the role of expertise and prey behavior on individual predator foraging specialization, and its potential fitness consequences.

The development of foraging strategies is crucial for young predators to reach adulthood and survive [@Phillips.etal2017;@Heithaus.etal2018]. Theory predicts that foraging specialization may emerge via learning, limitations in memorizing multiple complex hunting skills, and expertise [@Tinker.etal2009;@Dukas2019]. @Dukas2019 defines expertise as the characteristics, skills, and knowledge allowing individuals with extensive experience to outperform novices on complex tasks. This body of work suggests that the development of expertise is an optimizing process that promotes foraging specialization. Empirical studies on human and non-human hunters show that expertise optimizes the efficiency (e.g. search and handling times, return rates) of their foraging tactics potentially via associative images or reliance on prey and environmental cues [@Edwards.Jackson1994;@Morse2000;@macdonaldCrossculturalComparisonLearning2007;@Reid.etal2010;@Wilson-Rankin2015]. Such optimization may thus reinforce the use of the same tactic (i.e. specialization) if its success is constant each time a prey is encountered. It may also be costly to attempt different hunting tactics by trial and error when prey are scarce or highly unpredictable [@Dukas1998;@Estes.etal2003;@Mery.Burns2010]. An alternative mechanism is that expertise may offset the costs/risks of switching foraging tactics as individuals gain experience and information on their prey or their environment, leading to increased individual foraging flexibility [@Ishii.Shimada2010;@Mery.Burns2010;@Snell-Rood2013]. To develop proper responses to fluctuating resources, individuals would need to sample their environment broadly during an extensive period of time, followed by higher performance at later stages of development [reviewed in @Snell-Rood2013].

There is currently a lack of consensus on the fitness advantages of specialized vs flexible foraging. Importantly, we have little information on the ecological contexts - except for competition - that favor one strategy over the other. This is reflected in the literature showing contrasting results in the links between specialization and fitness. For instance, some studies report increasing benefits of specializing [@Patrick.Weimerskirch2014a;@vandenBosch.etal2019], some report that flexible foraging has greater benefits [@Paull.etal2012;@Manlick.etal2021], and others find equal benefits depending on timescales [@Woo.etal2008;@Potier.etal2015]. While quantifying the fitness consequences of specialization is a daunting task, in predator-prey systems, some clues indicate that fluctuations in the predictability of prey encounters throughout a predator's lifetime may be a key factor [@Weimerskirch2007;@Woo.etal2008;@Chang.etal2017;@Phillips.etal2017;@Courbin.etal2018]. The resource-predictability hypothesis advances that when resources are predictable, individual specialists should have higher delivery rates by reducing the energy and time required to search for and handle prey. In contrast, individual generalists should be advantaged when resources fluctuate, as fine adjustments to resources are key for a predator's success and survival [@Karkarey.etal2017;@Holm.etal2019;@Santoro.etal2019]. This hypothesis has however not been tested in systems where behavioral interactions are simultaneously monitored between individuals.

In this study, we analyze individual behavioral data from players in *Dead by Daylight* (*DBD*) to test how hunting expertise and prey behavior shape predator foraging specialization. *DBD* is an online videogame simulating a predator-prey interaction, where one predator player hunts four prey players in different virtual environments. The data grants a high degree of precision on the behavioral interaction as both the predator and the four prey were measured simultaneously in each trial. First, we investigate how predators developed their individual hunting expertise. We hypothesize that predators should differ in the development of their expertise because they encounter varying levels of difficulty with the prey that they hunt. Second, we test the hypothesis that individual variation in foraging behavior will change with expertise. If expertise reduces the costs of switching between hunting tactics, we predict that individual predators should become more flexible with time. Alternatively, if expertise enables the refinement of the hunting tactics, then individuals may instead specialize. Third, we evaluate how predator foraging specialization interacts with variation in prey behavior. Whether predators specialize or not with expertise may depend on the behavior of their prey, for instance, as it may be harder to specialize when encounters are less predictable. Lastly, if individuals differ in their degree of foraging specialization, then we expect specialist hunters to fare better when prey variability is lower, while flexible hunters should fare better when prey variability is higher.

\newpage
# MATERIALS AND METHODS {.unnumbered}

## Study system

*DBD* is a survival horror asymmetric multiplayer online game developed by Behavior Interactive Inc. In this game, players get to chose if they wish to play has a predator or a prey. The objective of the predator player is to search, stalk, and hunt the four prey players across a virtual environment. The objective of the four prey players is to forage for resources in the form of generators that once activated, will enable them to escape through one of two exit doors. A match ends when the predator kills all the prey available (i.e. that haven't escaped), or when the last prey left escapes the virtual environment. Each player, predator or prey, can chose an avatar with abilities that encourage specific play styles (e.g. bolder vs cautious prey, or ambush vs roaming predator). During our study period, the game offered 23 playable predator avatars. The virtual environment where matches take place is composed of fixed and procedurally generated items (e.g. vegetation, mazes, buildings) that constitute its structure. Some of these environments are larger than others, with varying levels or buildings. However, we have previously shown that they generally do not influence the behavior nor hunting success of the predator, which is probably due to a game feature enabling predators to have visual cues on the generators [@FraserFranco.etal2022]. There were 35 virtual game environments available for play during our study period.

Individual predators in *DBD* display variation in the hunting strategies that they use. Some specialize at hunting at high speeds and covering space in the environment, while others prefer to stalk and ambush their prey [@FraserFranco.etal2022]. They also display flexibility in the use of these tactics by switching between them across matches. The expression of these tactics and their success is highly dependent on the behavior of the prey group. The prey need to forage for resources while maintaining a high level of attention on the predator cues to avoid being detected and chased. Some prey contribute to the group's success by healing or helping others escape the predator, while other self-centered individuals play alone and attempt to escape by themselves [@Cere.etal2021]. Predators must be wary of how the prey behave to chose the right way to capture them. Thus, *DBD* simulates a highly dynamic system where both predators and prey must constantly adapt to each other to be successful.

## Data collection

We were provided with raw datasets from the videogame company that spanned a period of 6 months of matches recorded for each player.

# (mean = `r toString(format(round(mean(data$space_covered_rate, na.rm = TRUE), digits = 2) (pour écrire dans le texte)

## Behaviors

## Statistical analyses

### Software and computer components

### Effect of experience on hunting success

### Foraging tactics at each level of experience

\newpage
# RESULTS {.unnumbered}

## Effect of experience on hunting success

Contrary to our predictions, individual predators did not differ in the development of their expertise, as the best model was the one with a general population trend with varying intercepts only (Table x.). Moreover, we did not detect a relationship between experience (i.e. the cumulative amount of matches played) and hunting success, such that players did not optimize their success with increasing experience as we expected (Figure 1 A). However, predators did differ in their average hunting success across the study period (Figure 1 B).

```{r out.width = "100%", fig.cap = "**Figure 1.** Results from the Bayesian generalized linear mixed model. (A) Fitted (solid) line displaying the posterior predicted values relating the predators' hunting success to their cumulative experience. Here, we transformed the hunting success and cumulative experience back to their original scale (i.e. the amount of prey captured and the number of matches played respectively). The band around the solid line represents the 95% credible intervals. (B) The estimated posterior predicted mean hunting success of each predator player. We display the hunting success back-transformed to its original scale."}
knitr::include_graphics(here::here("outputs", "05_outputs_figures", "05_figure1.png"))
```

blablabla figure 2

```{r out.width = "100%", fig.cap = "**Figure 2.** Coefficient of variation for the predator's speed and for the speed of the prey encountered at each level of experience. (A) Coefficient of variation for the among individual differences in average behavior. (B) Coefficient fo variation for the among individual differences in within individual variance (i.e. specialization)."}
knitr::include_graphics(here::here("outputs", "05_outputs_figures", "05_figure2.png"))
```

blablabla figure 3

```{r out.width = "100%", fig.cap = "**Figure 3.** Title (A) Novice. (B) Advanced."}
knitr::include_graphics(here::here("outputs", "05_outputs_figures", "05_figure3.png"))
```

blablabla figure 4

```{r out.width = "100%", fig.cap = "**Figure 4.** Title (A) and (B) Novice. (C) and (D) Advanced."}
knitr::include_graphics(here::here("outputs", "05_outputs_figures", "05_figure4.png"))
```

\newpage
# LITTERATURE CITED {.unnumbered}

<div id="refs"></div>