### 2023-01-10 manuscript/Chapter2-SupInfo.Rmd
updated styles, ms, supmat, fig S1 S2, and fig1 code

---
title: "Hunting experience and prey variability jointly shape individual foraging specialisation in a predator-prey videogame"
output: 
  word_document:
    reference_docx: Chapter2-styles.docx
    fig_caption: yes
bibliography: references.bib
csl: e-letters.csl
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(data.table)
library(brms)
library(flextable)
library(officer)

```

need to talk about priors here for the gamms etc

```{r figS1, out.width = "100%", fig.cap = "**Figure S1.** Among individual differences in the development of hunting expertise. The predators' hunting success (i.e. the probability of capturing the four prey) is on the y axis, and the predators' cumulative experience (i.e. the number of matches played prior to each observation) is on the x axis. Each fitted curve represents an individual predator. (A) Individuals with the greatest increase in hunting success with experience (B) Individuals with the greatest decrease in hunting success with experience."}

knitr::include_graphics(here::here("outputs", "05_outputs_figures", "05_figureS1.png"))

```

```{r figS2, out.width = "100%", fig.cap = "**Figure S2.** Distribution of the intra individual behavioural variation (i.e. specialization) of predators with the greatest increase in hunting success (A and B) and greatest decrease (B and C) in hunting success when they were novices (A and C) and then (B and D) advanced hunters."}

knitr::include_graphics(here::here("outputs", "05_outputs_figures", "05_figureS2.png"))

```

```{r tableS1, eval = FALSE, ft.align = "left"}

tab1 <- readRDS(here::here("outputs", "05_outputs_tables", "05_table1.rds"))

set_caption(
  tab1,
  caption = "Posterior mean values, standard errors, and 95% credible intervals for the GAMM with a common global smoother. The values are presented on the logit scale.",
  style = "Table Caption"
)

```