### 2023-02-15 manuscript/Chapter2-SupInfo.Rmd
update supmat

---
title: |
     Online supporting information for \
     Hunting experience shapes individual foraging specialisation and predator-prey interactions in an online videogame

output: 
  word_document:
    reference_docx: Chapter2-styles.docx
    fig_caption: yes
bibliography: references.bib
csl: e-letters.csl
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(data.table)
library(brms)
library(flextable)
library(officer)
library(tidyr)

```

\newpage
# Materials and methods

## Statistical analyses

### Prior specifications

need to talk about priors here for the gamms etc

### Variance partitionning

We calculated the intra-class correlation coefficients ($ICC$, i.e. repeatability) of each random effect on the mean part of the predator's speed and prey speed to estimate the percentage of variation in speed attributed to differences among environments, avatars, and individuals. We also report the coefficient of variation in the mean (termed $CV_I$) and dispersion (termed $CV_p$) part of the predator and prey's speed, at each stage of experience, to report standardized effect sizes of individual variation for comparison across traits and studies [@cleasbyQuantifyingPredictabilityBehaviour2015;@Holtmann.etal2017;@Dingemanse.etal2022]. This also allowed us to evaluate whether predators displayed increasing/decreasing differences in specialization at each stage of experience.

\newpage
# Results

```{r figS1, out.width = "100%", fig.cap = "**Figure S1.** Among individual differences in the development of hunting expertise. The predators' hunting success (i.e. the probability of capturing the four prey) is on the y axis, and the predators' cumulative experience (i.e. the number of matches played prior to each observation) is on the x axis. Each fitted curve represents an individual predator. (A) Individuals with the greatest increase in hunting success with experience (B) Individuals with the greatest decrease in hunting success with experience."}

knitr::include_graphics(here::here("outputs", "05_outputs_figures", "05_figureS1.png"))

```

```{r figS2, out.width = "100%", fig.cap = "**Figure S2.** Distribution of the intra individual behavioural variation (i.e. specialization) of predators with the greatest increase in hunting success (A and B) and greatest decrease (C and D) in hunting success when they were novices (A and C) and then (B and D) advanced hunters."}

knitr::include_graphics(here::here("outputs", "05_outputs_figures", "05_figureS2.png"))

```

```{r figS4, out.width = "100%", fig.cap = "**Figure S4.** Posterior median differences and 95% credible intervals between the parameter values of each predator experience level predicted by the MDHGLM. The test is displayed on the y axis (i.e. whether the group of interest has either a greater or smaller value), and the parameter value is displayed on the x axis. Each panel represent the experience level groups that are being compared."}

knitr::include_graphics(here::here("outputs", "05_outputs_figures", "05_figureS4.png"))

```